---
title: "R Notebook"
output: html_notebook
---
############################################################
## 0) LIBRARIES + DATASETS
############################################################

```{r}
# Core
library(dplyr)
library(ggplot2)

theme_set(
  theme_bw() +          
  theme(
    axis.text    = element_text(size = 12),
    axis.title   = element_text(size = 14),
    legend.text  = element_text(size = 12),
    plot.title   = element_blank()          # removes all titles globally
  )
)
  
library(stringr)

# Models
library(lme4)
library(lmerTest)

# Diagnostics
library(car)          
library(performance)  

# Effects / predictions
library(ggeffects)
library(marginaleffects)

```


```{r}
df <- read.csv("Official_df_ideologyvspopulism.csv") %>%
  rename(term_id = term)

```

```{r}
df_crisis_examples <- df %>%
  mutate(
    Populism = as.numeric(Populism),
    crisis_systemic = as.numeric(crisis_systemic),
    crisis_security = as.numeric(crisis_security),
    crisis_healthhumanitarian = as.numeric(crisis_healthhumanitarian),
    crisis_damage = as.numeric(crisis_damage),
    crisis_economical = as.numeric(crisis_economical),
    crisis_natural_disasters = as.numeric(crisis_natural_disasters)
  ) %>%
  filter(
    !is.na(text),
    !is.na(Populism)
  )

# Helper function for z-scoring
z <- function(x) {
  as.numeric(scale(x))
}

# 1. Populist crisis appeal:
# high systemic crisis + high security crisis,
# low health/humanitarian + low damage crisis language
example_populist_crisis <- df_crisis_examples %>%
  filter(Populism == 1) %>%
  mutate(
    example_score =
      z(crisis_systemic) +
      z(crisis_security) -
      z(crisis_healthhumanitarian) -
      z(crisis_damage)
  ) %>%
  arrange(desc(example_score)) %>%
  slice(1)

# 2. Non-populist crisis appeal:
# high health/humanitarian + high damage crisis language,
# low systemic/security crisis language
example_nonpopulist_crisis <- df_crisis_examples %>%
  filter(Populism == 0) %>%
  mutate(
    example_score =
      z(crisis_healthhumanitarian) +
      z(crisis_damage) -
      z(crisis_systemic) -
      z(crisis_security)
  ) %>%
  arrange(desc(example_score)) %>%
  slice(1)

print_crisis_example <- function(x, title) {
  cat("\n", strrep("=", 100), "\n")
  cat(title, "\n")
  cat(strrep("=", 100), "\n")
  
  cat("Speech ID:", x$Speech_id, "\n")
  cat("Party:", x$party, "\n")
  cat("Person:", x$person, "\n")
  cat("Date:", as.character(x$date), "\n")
  cat("Populism:", x$Populism, "\n")
  cat("Example score:", x$example_score, "\n\n")
  
  cat("Systemic crisis:", x$crisis_systemic, "\n")
  cat("Security crisis:", x$crisis_security, "\n")
  cat("Health/humanitarian crisis:", x$crisis_healthhumanitarian, "\n")
  cat("Damage crisis:", x$crisis_damage, "\n")
  cat("Economic crisis:", x$crisis_economical, "\n")
  cat("Natural disasters:", x$crisis_natural_disasters, "\n\n")
  
  cat("TEXT:\n")
  cat(x$text, "\n")
}

print_crisis_example(
  example_populist_crisis,
  "POPULIST CRISIS APPEAL: HIGH SYSTEMIC + SECURITY CRISIS, LOW HEALTH/DAMAGE LANGUAGE"
)

print_crisis_example(
  example_nonpopulist_crisis,
  "NON-POPULIST CRISIS APPEAL: HIGH HEALTH/DAMAGE CRISIS, LOW SYSTEMIC/SECURITY LANGUAGE"
)
```




```{r}

table_populism <- df %>%
  filter(Populism == 1) %>%
  count(election_year, party, individual, name = "n_populist_paragraphs") %>%
  arrange(election_year, party, desc(n_populist_paragraphs))
```


```{r}
vars_to_scale <- c(
  "rile",
  "intensity_left",
  "intensity_right",
  "crisis",
  "real_crisis", 
  "constructed_crisis",
  "crisis_damage",
  "crisis_economical",
  "crisis_general",
  "crisis_healthhumanitarian",
  "crisis_natural_disasters",
  "crisis_security",
  "crisis_systemic",
  "formality",
  "scapegoating",
  "scapegoating_implicit",
  "scapegoating_victimhood",
  "scapegoating_victim_us",
  "scapegoating_victim_me",
  "scapegoating_them",
  "sentiment_negtw",
  "sentiment_postw",
  "sentiment_tottw",
  "sentiment_valtw",
  "sentiment_neg",
  "sentiment_pos",
  "sentiment_tot",
  "sentiment_val",
  "them_use",
  "us_use",
  "vulgarity"
)

df[vars_to_scale] <- lapply(df[vars_to_scale], function(x) as.numeric(scale(x)))

```



############################################################
## 2) DATA PREP FOR ALL INDIVIDUALS (BALANCING)
############################################################


```{r}
set.seed(42)

df_balanced <- df %>%
  group_by(individual) %>%
  group_modify(~ {
    n_pop <- sum(.x$Populism == 1)
    df_pop <- filter(.x, Populism == 1)
    df_non <- filter(.x, Populism == 0)

    if (nrow(df_non) >= n_pop) {
      bind_rows(df_pop, slice_sample(df_non, n = n_pop))
    } else {
      .x
    }
  }) %>%
  ungroup()

```

############################################################
## 3) BAD MANNERS — Formality
############################################################


```{r}
m_formality <- lmer(
  formality ~ Populism + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced
)

summary(m_formality)

```

```{r}
m_formality_pop <- glmer(
  Populism ~ formality + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_formality_pop)

```

```{r}
pred_formality <- ggpredict(m_formality_pop, terms = "formality [all]")

p <- ggplot(pred_formality, aes(x = x, y = predicted)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .15) +
  scale_y_continuous(trans = "logit") +
  labs(x = "Formality", y = "Predicted probability of Populism") +
  theme_minimal()

ggsave("predicted_populism_by_formality.png", plot = p, width = 7, height = 5, dpi = 300)

```

############################################################
## 3) BAD MANNERS — Profanity (vulgarity)
############################################################


```{r}
m_vulgarity <- lmer(
  vulgarity ~ Populism + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced
)

summary(m_vulgarity)

```


```{r}
m_vulgarity_pop <- glmer(
  Populism ~ vulgarity + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_vulgarity_pop)

```

```{r}
pred_profanity <- ggpredict(m_vulgarity_pop, terms = "vulgarity [all]")

p <- ggplot(pred_profanity, aes(x = x, y = predicted)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .15) +
  scale_y_continuous(trans = "logit") +
  labs(x = "Profanity", y = "Predicted probability of Populism") +
  theme_minimal()

ggsave("predicted_populism_by_vulgarity.png", plot = p, width = 7, height = 5, dpi = 300)

```

############################################################
## 3) BAD MANNERS — Sentiment
############################################################


```{r}
# sentiment_label model 
m_sentimentvt_pop <- glmer(
  Populism ~ sentiment_label + prior_president + rile + party +
    factor(election_year) + (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_sentimentvt_pop)

```


```{r}
m_sentimentvt_pop1 <- glmer(
  Populism ~ sentiment_valtw + prior_president + rile + party +
    factor(election_year) + (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_sentimentvt_pop1)

```

```{r}
pred_sentiment <- ggpredict(m_sentimentvt_pop1, terms = "sentiment_valtw [all]")

p <- ggplot(pred_sentiment, aes(x = x, y = predicted)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .15) +
  scale_y_continuous(trans = "logit") +
  labs(x = "Sentiment Valence", y = "Predicted probability of Populism") +
  theme_minimal()

ggsave("predicted_populism_by_sentiment_valtw.png", plot = p, width = 7, height = 5, dpi = 300)

```


```{r}
sent_vars <- c("sentiment_valtw", "sentiment_tottw", "sentiment_neg", "sentiment_pos")

m_sent_lmm <- list()

for (v in sent_vars) {
  cat("\n\n========================================\n")
  cat("LMM: Outcome =", v, "\n")
  cat("========================================\n")

  f <- as.formula(
    paste0(
      v,
      " ~ Populism + rile + prior_president + party + factor(election_year) + (1 | individual)"
    )
  )

  m_sent_lmm[[v]] <- lmer(f, data = df_balanced)
  print(summary(m_sent_lmm[[v]]))
}



m_sent_glmm <- list()

for (v in sent_vars) {
  cat("\n\n========================================\n")
  cat("GLMM: Predictor =", v, "\n")
  cat("========================================\n")

  f <- as.formula(
    paste0(
      "Populism ~ ",
      v,
      " + rile + party + prior_president + factor(election_year) + (1 | individual)"
    )
  )

  m_sent_glmm[[v]] <- glmer(
    f,
    data = df_balanced,
    family = binomial,
    control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 5e5))
  )

  print(summary(m_sent_glmm[[v]]))
}

```




############################################################
## 3) BAD MANNERS — Tot
############################################################


```{r}
m_badmanners <- glmer(
  Populism ~ formality + vulgarity + sentiment_valtw + prior_president + rile + party +
    factor(election_year) + (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_badmanners)

```


```{r}
check_collinearity(m_badmanners)

```


```{r}
m_badmanners_form <- lmer(
  formality ~ Populism + vulgarity + sentiment_valtw + prior_president + rile + party +
    factor(election_year) + (1 | individual),
  data = df_balanced
)

summary(m_badmanners_form)

```
```{r}
m_badmanners_prof <- lmer(
   vulgarity~ Populism + formality + sentiment_valtw + prior_president + rile + party +
    factor(election_year) + (1 | individual),
  data = df_balanced
)

summary(m_badmanners_prof)

```
```{r}
m_badmanners_prof <- lmer(
   sentiment_valtw~ Populism + formality + vulgarity + prior_president + rile + party +
    factor(election_year) + (1 | individual),
  data = df_balanced
)

summary(m_badmanners_prof)

```

############################################################
## 4) US VS THEM — Us vs Them (them_use)
############################################################


```{r}
m_them <- lmer(
  them_use ~ Populism + prior_president + rile + party +
    factor(election_year) + (1 | individual),
  data = df_balanced
)

summary(m_them)

```

```{r}
m_them_pop <- glmer(
  Populism ~ them_use + prior_president + rile + party +
    factor(election_year) + (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_them_pop)

```


############################################################
## 4) US VS THEM — Us vs Them (Us_use)
############################################################


```{r}
m_us <- lmer(
  us_use ~ Populism + prior_president + rile + party +
    factor(election_year) + (1 | individual),
  data = df_balanced
)

summary(m_us)
```

```{r}
m_us_pop <- glmer(
  Populism ~ us_use + prior_president + rile + party +
    factor(election_year) + (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_us_pop)
```

```{r}
pred_ususe <- ggpredict(m_us_pop, terms = "us_use [all]")

p <- ggplot(pred_ususe, aes(x = x, y = predicted)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .15) +
  scale_y_continuous(trans = "logit") +
  labs(x = "People-Centric Rhetoric", y = "Predicted probability of Populism") +
  theme_minimal()

ggsave("predicted_populism_by_us_use.png", plot = p, width = 7, height = 5, dpi = 300)
```



############################################################
## 4) US VS THEM — Us vs Them (tot_use)
############################################################

```{r}
m_us_them_pop <- glmer(
  Populism ~ us_use + them_use + prior_president + rile + party +
    factor(election_year) + (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_us_them_pop)
```
```{r}
check_collinearity(m_us_them_pop)
```

```{r}
library(emmeans)

grid <- expand.grid(
  us_use = seq(min(df$us_use), max(df$us_use), length = 50),
  them_use = seq(min(df$them_use), max(df$them_use), length = 50),
  rile = mean(df$rile),
  prior_president = 0,
  party = "dem",
  election_year = 2000
)

grid$pred <- predict(
  m_us_them_pop, 
  newdata = grid, 
  type = "response", 
  re.form = NA
)

p<-ggplot(grid, aes(us_use, them_use, fill = pred)) +
  geom_tile() +
  scale_fill_viridis_c(name = "P(Populism)") +
  labs(
    x = "People-Centric Rhetoric",
    y = "Anti-Elite Rhetoric"
  ) +
  theme_minimal()
ggsave("predicted_populism_by_usthem_use.png", plot = p, width = 7, height = 5, dpi = 300)
```


############################################################
## 4) US VS THEM — Scapegoating - blaming, causality
############################################################

```{r}
m_scapegoating <- lmer(
  scapegoating ~ Populism + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced
)

summary(m_scapegoating)

```

```{r}
m_scapegoating_pop <- glmer(
  Populism ~ scapegoating + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_scapegoating_pop)

```


############################################################
## 4) US VS THEM — Scapegoating - sentence_expanded
############################################################

```{r}
m_scapegoating_expanding <- lmer(
  scapegoating_implicit ~ Populism + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced
)

summary(m_scapegoating_expanding)

```

```{r}
m_scapegoating_expanding_pop <- glmer(
  Populism ~ scapegoating_implicit + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_scapegoating_expanding_pop)

```
############################################################
## 4) US VS THEM — Scapegoating - sentence_them
############################################################

```{r}
m_scapegoating_them <- lmer(
  scapegoating_them ~ Populism + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced
)

summary(m_scapegoating_them)

```

```{r}
m_scapegoating_pop_them <- glmer(
  Populism ~ scapegoating_them + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_scapegoating_pop_them)

```

############################################################
## 4) US VS THEM — Victimism - sentence_general
############################################################

```{r}
m_scapegoating_victimhood <- lmer(
  scapegoating_victimhood ~ Populism + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced
)

summary(m_scapegoating_victimhood)

```

```{r}
m_scapegoating_victimhood_pop <- glmer(
  Populism ~ scapegoating_victimhood + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_scapegoating_victimhood_pop)

```

############################################################
## 4) US VS THEM — Victimism - sentence_us
############################################################

```{r}
m_scapegoating_us <- lmer(
  scapegoating_victim_us ~ Populism + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced
)

summary(m_scapegoating_us)

```

```{r}
m_scapegoating_pop_us <- glmer(
  Populism ~ scapegoating_victim_us + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_scapegoating_pop_us)

```

############################################################
## 4) US VS THEM — Victimism - sentence_me
############################################################

```{r}
m_scapegoating_me <- lmer(
  scapegoating_victim_me ~ Populism + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced
)

summary(m_scapegoating_me)

```

```{r}
m_scapegoating_pop_me <- glmer(
  Populism ~ scapegoating_victim_me + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_scapegoating_pop_me)

```


############################################################
## 4) US VS THEM — tot - scapegoating
############################################################

```{r}
m_scapegoating_tot <- glmer(
  Populism ~ scapegoating + scapegoating_implicit + scapegoating_them + scapegoating_victimhood + scapegoating_victim_us + scapegoating_victim_me + rile + prior_president + party + factor(election_year) + (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_scapegoating_tot)

```

```{r}
check_collinearity(m_scapegoating_tot)
```
```{r}
ame <- avg_slopes(
  m_scapegoating_tot,
  variables = c("scapegoating",
                "scapegoating_implicit",
                "scapegoating_them",
                "scapegoating_victimhood",
                "scapegoating_victim_us",
                "scapegoating_victim_me"),
  type = "response"
)

ame_pp <- ame |>
  mutate(
    estimate = estimate * 100,
    conf.low = conf.low * 100,
    conf.high = conf.high * 100,
    term = dplyr::recode(term,
      scapegoating = "Blame/Causality",
      scapegoating_implicit = "Generic Scapegoating",
      scapegoating_them = "Out-group Scapegoating",
      scapegoating_victimhood = "Generic Victimhood",
      scapegoating_victim_us = "In-Group Victimhood",
      scapegoating_victim_me = "Self-Victimhood")
  )

ame_pp$term <- factor(ame_pp$term, levels = rev(ame_pp$term))

p <- ggplot(ame_pp, aes(x = estimate, y = term)) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_point(size = 2) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  labs(x = "Average Marginal Effect on Probability of Populism", y = NULL) +
  theme_minimal()

ggsave("predicted_populism_by_scapegoating.png", plot = p, width = 7, height = 5, dpi = 300)
```


############################################################
## 5) CRISES
############################################################


```{r}
m_crisis <- lmer(
  crisis ~ Populism + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced
)

summary(m_crisis)

```

```{r}
m_crisis_pop <- glmer(
  Populism ~ crisis + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "nloptwrap", optCtrl = list(maxfun = 2e5))
)

summary(m_crisis_pop)

```
############################################################
## 5) CRISES - Real
############################################################


```{r}
m_crisis_real <- lmer(
  real_crisis ~ Populism + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced
)

summary(m_crisis_real)

```

```{r}
m_crisis_pop_real <- glmer(
  Populism ~ real_crisis + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_crisis_pop_real)

```

############################################################
## 5) CRISES - Constructed
############################################################


```{r}
m_crisis_Constructed <- lmer(
  constructed_crisis ~ Populism + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced
)

summary(m_crisis_Constructed)

```

```{r}
m_crisis_pop_Constructed <- glmer(
  Populism ~ constructed_crisis + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "nloptwrap", optCtrl = list(maxfun = 2e5))
)

summary(m_crisis_pop_Constructed)

```
############################################################
## 5) CRISES - categories
############################################################

```{r}
# Crisis category variables you want separate regressions for
crisis_vars <- c(
  "crisis_damage",
  "crisis_economical",
  "crisis_general",
  "crisis_healthhumanitarian",
  "crisis_natural_disasters",
  "crisis_security",
  "crisis_systemic"
)

# Store models here (optional but recommended)
m_crisis_cat <- list()
m_crisis_pop_cat <- list()

for (v in crisis_vars) {

  # 1) Outcome model: category score ~ Populism + controls
  f1 <- as.formula(paste0(
    v, " ~ Populism + rile + prior_president + party + factor(election_year) + (1 | individual)"
  ))

  m_crisis_cat[[v]] <- lmer(f1, data = df_balanced)

  cat("\n############################################################\n")
  cat("## CRISIS CATEGORY (lmer): ", v, "\n")
  cat("############################################################\n\n")
  print(summary(m_crisis_cat[[v]]))


  # 2) Populism model: Populism ~ category score + controls
  f2 <- as.formula(paste0(
    "Populism ~ ", v, " + rile + prior_president + party + factor(election_year) + (1 | individual)"
  ))

  m_crisis_pop_cat[[v]] <- glmer(
    f2,
    data = df_balanced,
    family = binomial,
    control = glmerControl(
      optimizer = "bobyqa",
      optCtrl = list(maxfun = 2e5)
    )
  )

  cat("\n############################################################\n")
  cat("## CRISIS CATEGORY (glmer): Populism ~ ", v, "\n")
  cat("############################################################\n\n")
  print(summary(m_crisis_pop_cat[[v]]))
}
```


############################################################
## 5) CRISES - tot
############################################################


```{r}
m_crisis_tot_pop <- glmer(
  Populism ~ crisis + real_crisis + constructed_crisis + rile + prior_president + party +
    factor(election_year) + (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_crisis_tot_pop)

```

```{r}
m_crisis_tot_pop_cat <- glmer(
  Populism ~ 
    crisis_damage + crisis_economical  + crisis_general +
    crisis_healthhumanitarian + crisis_natural_disasters +
    crisis_security + crisis_systemic +
    rile + prior_president + party + (1|election_year) +
    (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_crisis_tot_pop_cat)


```

```{r}
check_collinearity(m_crisis_tot_pop_cat)
```
```{r}
ame_crisis <- avg_slopes(
  m_crisis_tot_pop_cat,
  variables = c(
    "crisis_damage",
    "crisis_economical",
    "crisis_general",
    "crisis_healthhumanitarian",
    "crisis_natural_disasters",
    "crisis_security",
    "crisis_systemic"
  ),
  type = "response"
)

ame_crisis_pp <- ame_crisis %>%
  mutate(
    estimate  = estimate * 100,
    conf.low  = conf.low * 100,
    conf.high = conf.high * 100,
    term = dplyr::recode(term,
      crisis_damage = "Damage",
      crisis_economical = "Economic",
      crisis_general = "Generic",
      crisis_healthhumanitarian = "Health/Humanitarian",
      crisis_natural_disasters = "Natural disasters",
      crisis_security = "Security",
      crisis_systemic = "Systemic"
    )
  )

ame_crisis_pp$term <- factor(ame_crisis_pp$term, levels = rev(ame_crisis_pp$term))


# Plot
p <- ggplot(ame_crisis_pp, aes(x = estimate, y = term)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(size = 2) +
  geom_errorbarh(
    aes(xmin = conf.low, xmax = conf.high),
    height = 0.2
  ) +
  labs(
    x = "Average Marginal Effect on Probability of Populism",
    y = NULL
  ) +
  theme_minimal()

ggsave(
  "predicted_populism_by_crisis_categories.png",
  plot = p,
  width = 7,
  height = 5,
  dpi = 300
)
```


############################################################
## 6) TOTAL INFLUENCE ON POPULISM - all individuals
##    (cumulative effects of the final regressions)
############################################################



```{r}
m_tot_full <- glmer(
  Populism ~ formality + vulgarity + sentiment_valtw +
     crisis_damage + crisis_economical  + crisis_general +
    crisis_healthhumanitarian + crisis_natural_disasters +
    crisis_security + crisis_systemic +
    us_use + them_use +
    scapegoating  + scapegoating_implicit  + scapegoating_victimhood + scapegoating_victim_me +
    rile + party + prior_president + 
    (1|election_year) + (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_tot_full)

```
```{r}
check_collinearity(m_tot_full)
```

```{r}
m_tot_full_int <- glmer(
  Populism ~ formality + vulgarity + sentiment_valtw +
     crisis_damage + crisis_economical  + crisis_general +
    crisis_healthhumanitarian + crisis_natural_disasters +
    crisis_security + crisis_systemic +
    us_use + them_use +
    scapegoating  + scapegoating_implicit  + scapegoating_victimhood + scapegoating_victim_me +
     intensity_left + intensity_right + party + prior_president + 
    (1|election_year) + (1 | individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_tot_full_int)
```

```{r}
check_collinearity(m_tot_full_int)
```



```{r}
m_ideology_tot_full <- lmer(
  rile ~ formality + vulgarity + sentiment_valtw +
     crisis_damage + crisis_economical  + crisis_general +
    crisis_healthhumanitarian + crisis_natural_disasters +
    crisis_security + crisis_systemic +
    us_use + them_use +
    scapegoating  + scapegoating_implicit  + scapegoating_victimhood + scapegoating_victim_me +
    Populism + party + prior_president + 
    (1|election_year) + (1 | individual),
  data = df_balanced
)

summary(m_ideology_tot_full)
```
```{r}
check_collinearity(m_ideology_tot_full)
```

```{r}
m_ideology_tot_intensel <- lmer(
  intensity_left ~ formality + vulgarity + sentiment_valtw +
     crisis_damage + crisis_economical  + crisis_general +
    crisis_healthhumanitarian + crisis_natural_disasters +
    crisis_security + crisis_systemic +
    us_use + them_use +
    scapegoating  + scapegoating_implicit  + scapegoating_victimhood + scapegoating_victim_me + intensity_right +
    Populism + party + prior_president + 
    (1|election_year) + (1 | individual),
  data = df_balanced
)

summary(m_ideology_tot_intensel)
```

```{r}
check_collinearity(m_ideology_tot_intensel)
```


```{r}
m_ideology_tot_intenser <- lmer(
  intensity_right ~ formality + vulgarity + sentiment_valtw +
     crisis_damage + crisis_economical  + crisis_general +
    crisis_healthhumanitarian + crisis_natural_disasters +
    crisis_security + crisis_systemic +
    us_use + them_use +
    scapegoating  + scapegoating_implicit  + scapegoating_victimhood + scapegoating_victim_me + intensity_left +
    Populism + party + prior_president + 
    (1|election_year) + (1 | individual),
  data = df_balanced
)

summary(m_ideology_tot_intenser)
```

```{r}
check_collinearity(m_ideology_tot_intenser)
```

```{r}
m_tot_full_pres <- glmer(
   prior_president ~ formality + vulgarity + sentiment_valtw +
     crisis_damage + crisis_economical  + crisis_general +
    crisis_healthhumanitarian + crisis_natural_disasters +
    crisis_security + crisis_systemic +
    us_use + them_use +
    scapegoating  + scapegoating_implicit  + scapegoating_victimhood + scapegoating_victim_me +
    rile + party + Populism + 
    (1|individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_tot_full_pres)
```

```{r}
check_collinearity(m_tot_full_pres)
```
```{r}
m_tot_full_pres_intense <- glmer(
   prior_president ~ formality + vulgarity + sentiment_valtw +
     crisis_damage + crisis_economical  + crisis_general +
    crisis_healthhumanitarian + crisis_natural_disasters +
    crisis_security + crisis_systemic +
    us_use + them_use +
    scapegoating  + scapegoating_implicit  + scapegoating_victimhood + scapegoating_victim_me +
    intensity_left + intensity_right + party + Populism + 
    (1|individual),
  data = df_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_tot_full_pres_intense)
```

```{r}
check_collinearity(m_tot_full_pres_intense)
```




############################################################
## 6) TOTAL INFLUENCE ON POPULISM - only prior_presidents
##    (cumulative effects of the final regressions)
############################################################

############################################################
## IDEOLOGY (df_switchers + calculation)
############################################################


```{r}
switchers <- c(
  "Bush_GeorgeHW", "Carter_Jimmy", "Clinton_Bill", "Eisenhower_Dwight",
  "Nixon_Richard", "Obama_Barack", "Reagan_Ronald", "Trump_Donald"
)

df_switchers <- df %>%
  filter(individual %in% switchers)

```

```{r}
set.seed(42)

df_switchers_balanced <- df_switchers %>%
  group_by(individual) %>%
  group_modify(~ {
    n_pop <- sum(.x$Populism == 1)
    df_pop <- filter(.x, Populism == 1)
    df_non <- filter(.x, Populism == 0)

    if (nrow(df_non) >= n_pop) {
      bind_rows(df_pop, slice_sample(df_non, n = n_pop))
    } else {
      .x
    }
  }) %>%
  ungroup()

```


```{r}
df_switchers_balanced %>%
  group_by(individual) %>%
  summarise(
    n_pop = sum(Populism == 1),
    n_non = sum(Populism == 0)
  )

```



############################################################
## 1B) Populism vs. Ideology (switchers) — regressions
############################################################


```{r}
# Populism vs. Ideology, populism
m_pop_logit <- glmer(
  Populism ~ rile + prior_president + party + (1|individual),
  data = df_switchers_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_pop_logit)

```

```{r}
check_collinearity(m_pop_logit)
```

```{r}
# Populism vs. Ideology, populism
m_pop_logit <- glmer(
  Populism ~ intensity_left + intensity_right + prior_president + party + (1|individual),
  data = df_switchers_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_pop_logit)

```


```{r}
# Populism vs. Ideology, ideology better
m_rile_lm_party <- lm(
  rile ~ party,
  data = df_switchers_balanced
)

summary(m_rile_lm_party)

```


```{r}
# Populism vs. Ideology, ideology better
m_rile_lmm_2 <- lmer(
  rile ~ Populism + prior_president + party + (1 | individual),
  data = df_switchers_balanced
)

summary(m_rile_lmm_2)

```


```{r}
m_tot_pres_switchers <- glm(
  prior_president ~ Populism + rile + party,
  data = df_switchers_balanced,
  family = binomial
)

summary(m_tot_pres_switchers)

```

```{r}
m_tot_full <- glmer(
  Populism ~ formality + vulgarity + sentiment_valtw +
     crisis_damage + crisis_economical  + crisis_general +
    crisis_healthhumanitarian + crisis_natural_disasters +
    crisis_security + crisis_systemic +
    us_use + them_use +
    scapegoating  + scapegoating_implicit  + scapegoating_victimhood + scapegoating_victim_me +
    rile + party + prior_president + 
    (1|individual),
  data = df_switchers_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_tot_full)

```


```{r}
check_collinearity(m_tot_full)
```


```{r}
m_tot_full_intensity <- glmer(
  Populism ~ formality + vulgarity + sentiment_valtw +
     crisis_damage + crisis_economical  + crisis_general +
    crisis_healthhumanitarian + crisis_natural_disasters +
    crisis_security + crisis_systemic +
    us_use + them_use +
    scapegoating  + scapegoating_implicit  + scapegoating_victimhood + scapegoating_victim_me +
    intensity_left + intensity_right + party + prior_president + 
    (1|individual),
  data = df_switchers_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(m_tot_full_intensity)

```
```{r}
check_collinearity(m_tot_full_intensity)
```


```{r}
m_ideology_tot_full <- lmer(
  rile ~ formality + vulgarity + sentiment_valtw +
     crisis_damage + crisis_economical  + crisis_general +
    crisis_healthhumanitarian + crisis_natural_disasters +
    crisis_security + crisis_systemic +
    us_use + them_use +
    scapegoating  + scapegoating_implicit  + scapegoating_victimhood + scapegoating_victim_me +
    Populism + party + prior_president + (1|individual),
  data = df_switchers_balanced
)

summary(m_ideology_tot_full)
```


```{r}
check_collinearity(m_ideology_tot_full)
```

```{r}
m_ideology_tot_left <- lmer(
  intensity_left ~ formality + vulgarity + sentiment_valtw +
     crisis_damage + crisis_economical  + crisis_general +
    crisis_healthhumanitarian + crisis_natural_disasters +
    crisis_security + crisis_systemic +
    us_use + them_use +
    scapegoating  + scapegoating_implicit  + scapegoating_victimhood + scapegoating_victim_me +
    Populism + party + prior_president + intensity_right + (1|individual),
  data = df_switchers_balanced
)

summary(m_ideology_tot_left)
```



```{r}
m_ideology_tot_right <- lmer(
  intensity_right ~ formality + vulgarity + sentiment_valtw +
     crisis_damage + crisis_economical  + crisis_general +
    crisis_healthhumanitarian + crisis_natural_disasters +
    crisis_security + crisis_systemic +
    us_use + them_use +
    scapegoating  + scapegoating_implicit  + scapegoating_victimhood + scapegoating_victim_me +
    Populism + party + prior_president + intensity_left + (1|individual),
  data = df_switchers_balanced
)

summary(m_ideology_tot_right)
```

```{r}
m_tot_full_pres <- glmer(
   prior_president ~ formality + vulgarity + sentiment_valtw +
     crisis_damage + crisis_economical  + crisis_general +
    crisis_healthhumanitarian + crisis_natural_disasters +
    crisis_security + crisis_systemic +
    us_use + them_use +
    scapegoating  + scapegoating_implicit  + scapegoating_victimhood + scapegoating_victim_me +
    rile + party + Populism + (1|individual),
  data = df_switchers_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
  
)

summary(m_tot_full_pres)
```


```{r}
m_tot_full_pres_int <- glmer(
   prior_president ~ formality + vulgarity + sentiment_valtw +
     crisis_damage + crisis_economical  + crisis_general +
    crisis_healthhumanitarian + crisis_natural_disasters +
    crisis_security + crisis_systemic +
    us_use + them_use +
    scapegoating  + scapegoating_implicit  + scapegoating_victimhood + scapegoating_victim_me +
    intensity_left + intensity_right + party + Populism + (1|individual),
  data = df_switchers_balanced,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
  
)

summary(m_tot_full_pres_int)
```


